---
title: "fig7_8"
author: "izel"
date: "2025-05-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This code generates Figure 7 and 8 (Supplementary Figures)

# Load packages

```{r warning=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
library(patchwork) # combine plots with arithmetic operations
library(reshape2)
```

# S3 Fig (outbreak size by the age group of index case)

# outbreak size across all age groups

```{r}
outbreak_size_no_control <- function(data_list, p_asym) {
  
  outbreak_size_no_control <- list()
  
  for (i in 1:length(data_list)) {
    
    outbreak_size_no_control[[i]] <- data_list[[i]][[3]] %>%
      mutate(cases_grp = case_when(cases_with_control >= 0 & cases_with_control < 1 ~ "[0]",
                                   cases_with_control >= 1 & cases_with_control < 2 ~ "[1]",
                                   cases_with_control >= 2 & cases_with_control < 5 ~ "[2, 4]",
                                   cases_with_control >= 5 & cases_with_control < 10 ~ "[5, 9]",
                                   cases_with_control >= 10 & cases_with_control < 20 ~ "[10, 19]",
                                   cases_with_control >= 20 & cases_with_control < 50 ~ "[20, 49]",
                                   cases_with_control >= 50  ~ "[>=50]"))
  }
  
  outbreak_grp_no_control <- list()
  
  for (i in 1:length(outbreak_size_no_control)) {
    
    outbreak_grp_no_control[[i]] <- outbreak_size_no_control[[i]] %>%
      group_by(cases_grp) %>%
      tally() %>%
      ungroup()
  }
  
  for (i in 1:length(outbreak_grp_no_control)) {
    outbreak_grp_no_control[[i]] <- outbreak_grp_no_control[[i]] %>%
      mutate(prop = n/sum(n))
    
  }
  
  outbreak_grp_no_control[[1]]$R_0 <- 0
  outbreak_grp_no_control[[2]]$R_0 <- 0.1
  outbreak_grp_no_control[[3]]$R_0 <- 0.2
  outbreak_grp_no_control[[4]]$R_0 <- 0.3
  outbreak_grp_no_control[[5]]$R_0 <- 0.4
  outbreak_grp_no_control[[6]]$R_0 <- 0.5
  outbreak_grp_no_control[[7]]$R_0 <- 0.6
  outbreak_grp_no_control[[8]]$R_0 <- 0.7
  outbreak_grp_no_control[[9]]$R_0 <- 0.8
  outbreak_grp_no_control[[10]]$R_0 <- 0.9
  outbreak_grp_no_control[[11]]$R_0 <- 1
  outbreak_grp_no_control[[12]]$R_0 <- 1.1
  
  outbreak_no_control_df <- do.call(rbind, outbreak_grp_no_control)
  
  outbreak_no_control_df$cases_grp <- factor(outbreak_no_control_df$cases_grp, levels = c("[0]", "[1]",
                                                                                    "[2, 4]", "[5, 9]",
                                                                                    "[10, 19]", "[20, 49]",
                                                                                    "[>=50]"))
  
  plot <- ggplot(outbreak_no_control_df, aes(x = factor(R_0), y = prop, fill = cases_grp)) +
    geom_bar(stat = "identity") +
    labs(x = expression(R[0]), y = "Proportion of simulation", 
         title = "Across all age groups", fill = "Outbreak size \n(excluding index case)") +
    scale_fill_manual(values = c("[0]" = "red",
                                 "[1]" = "orange",
                                 "[2, 4]" = "yellow",
                                 "[5, 9]" = "lightgreen",
                                 "[10, 19]" = "green",
                                 "[20, 49]" = "blue",
                                 "[>=50]" = "purple")) +
    theme_minimal() +
    theme(title = element_text(size = 8),
          axis.title.y = element_text(size = 10),
        legend.title = element_text(size = 10),
        legend.box.margin = margin(0, 0, 0, -10),
        legend.text = element_text(size = 9),
        legend.key.size = unit(0.8, "lines")) +
    theme(plot.background = element_rect(fill = "white", colour = NA))
  
  return(list(plot = plot, outbreak_no_control_df = outbreak_no_control_df, outbreak_grp_no_control = outbreak_grp_no_control, outbreak_size_no_control = outbreak_size_no_control))
}
```

```{r}
list_4a_no_control <- list(res_04_0_0_500_0, res_04_0_01_500_0, res_04_0_02_500_0, res_04_0_03_500_0, res_04_0_04_500_0, res_04_0_05_500_0, res_04_0_06_500_0, res_04_0_07_500_0, res_04_0_08_500_0, res_04_0_09_500_0, res_04_0_1_500_0, res_04_0_11_500_0)
```

Figure 4a: no control

```{r}
fig_all_age <- outbreak_size_no_control(list_4a_no_control, 0.4)$plot
```

# age-specific outbreak size

Data preparation for plotting

```{r}
list_no_control <- list(res_04_0_0_500_0, res_04_0_01_500_0, res_04_0_02_500_0, res_04_0_03_500_0, res_04_0_04_500_0,
                    res_04_0_05_500_0, res_04_0_06_500_0, res_04_0_07_500_0, res_04_0_08_500_0, res_04_0_09_500_0, res_04_0_1_500_0,
                    res_04_0_11_500_0)

outcome_list <- list()

for (data in 1:length(list_no_control)) {
  
  outbreak_size_age <- list_no_control[[data]][[2]][, c(1:3)]
  
  outbreak_size_age <- outbreak_size_age %>%
    mutate(sec_cases_grp = case_when(secondary_cases >= 0 & secondary_cases < 1 ~ "[0]",
                                   secondary_cases >= 1 & secondary_cases < 2 ~ "[1]",
                                   secondary_cases >= 2 & secondary_cases < 5 ~ "[2, 4]",
                                   secondary_cases >= 5 & secondary_cases < 10 ~ "[5, 9]",
                                   secondary_cases >= 10 & secondary_cases < 20 ~ "[10, 19]",
                                   secondary_cases >= 20 & secondary_cases < 50 ~ "[20, 49]",
                                   secondary_cases >= 50  ~ "[>=50]"))
  
  outbreak_size_age <- outbreak_size_age %>%
    group_by(age_index_case, sec_cases_grp) %>%
    tally() %>%
    ungroup()
  
  
  levels <- c("[0]", "[1]", "[2, 4]", "[5, 9]", "[10, 19]", "[20, 49]", "[>=50]")

  outbreak_size_age <- outbreak_size_age %>%
    complete(age_index_case, sec_cases_grp = levels, fill = list(n = 0))
  
  for (i in 1:length(outbreak_size_age$age_index_case)) {
  
    if (i <= 7) {outbreak_size_age$prop[i] <- outbreak_size_age$n[i]/sum(outbreak_size_age$n[1:7])}
    else if (i <= 14) {outbreak_size_age$prop[i] <- outbreak_size_age$n[i]/sum(outbreak_size_age$n[8:14])}
    else if (i <= 21) {outbreak_size_age$prop[i] <- outbreak_size_age$n[i]/sum(outbreak_size_age$n[15:21])}
    else {outbreak_size_age$prop[i] <- outbreak_size_age$n[i]/sum(outbreak_size_age$n[22:28])}
  
  }

  outcome_list[[data]] <- outbreak_size_age
  
}

outcome_list[[1]]$R <- 0
outcome_list[[2]]$R <- 0.1
outcome_list[[3]]$R <- 0.2
outcome_list[[4]]$R <- 0.3
outcome_list[[5]]$R <- 0.4
outcome_list[[6]]$R <- 0.5
outcome_list[[7]]$R <- 0.6
outcome_list[[8]]$R <- 0.7
outcome_list[[9]]$R <- 0.8
outcome_list[[10]]$R <- 0.9
outcome_list[[11]]$R <- 1
outcome_list[[12]]$R <- 1.1

size_age_R <- rbind(outcome_list[[1]], outcome_list[[2]], outcome_list[[3]], outcome_list[[4]],
                    outcome_list[[5]], outcome_list[[6]], outcome_list[[7]], outcome_list[[8]],
                    outcome_list[[9]], outcome_list[[10]], outcome_list[[11]], outcome_list[[12]])

```

Generate plots

```{r}
size_age_R_20_44 <- size_age_R %>%
  filter(age_index_case == "20-44")

levels <- c("[0]", "[1]", "[2, 4]", "[5, 9]", "[10, 19]", "[20, 49]", "[>=50]")
  
size_age_R_20_44$sec_cases_grp <- factor(size_age_R_20_44$sec_cases_grp, levels = levels)

x_vals <- as.character(seq(0, 1.1, 0.1))
positions <- rep(c(-0.03, -0.075), length.out = length(x_vals))  # y-offset for label placement

label_df <- data.frame(
  R = x_vals,
  prop = positions
)


plot_20_44_base <- ggplot() +
  geom_bar(data = size_age_R_20_44, aes(x = factor(R), y = prop, fill = sec_cases_grp), stat = "identity") +
  labs(x = expression(R[0]), y = "Proportion of simulations", title = "Age group index: 19-44",
       fill = "number of cases") +
  scale_fill_manual(values = c("[0]" = "red",
                                 "[1]" = "orange",
                                 "[2, 4]" = "yellow",
                                 "[5, 9]" = "lightgreen",
                                 "[10, 19]" = "green",
                                 "[20, 49]" = "blue",
                                 "[>=50]" = "purple")) +
  scale_y_continuous(breaks = seq(0, 1, 0.25)) +
  theme_minimal()+
  theme(legend.position = "none") + 
  theme(title = element_text(size = 8),
        axis.title.y = element_text(size = 10),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y = element_text(size = 8)) +
  coord_cartesian(ylim = c(0, 1), clip = "off")

plot_20_44 <- plot_20_44_base +
  geom_text(data = label_df,
            aes(x = R, y = prop, label = R),
            size = 2.8,
            inherit.aes = FALSE) + 
  theme(
  axis.title.x = element_text(size = 10, margin = margin(t = 10))
)

###

size_age_R_45_59 <- size_age_R %>%
  filter(age_index_case == "45-59")

levels <- c("[0]", "[1]", "[2, 4]", "[5, 9]", "[10, 19]", "[20, 49]", "[>=50]")
  
size_age_R_45_59$sec_cases_grp <- factor(size_age_R_45_59$sec_cases_grp, levels = levels)

plot_45_59_base <- ggplot() +
  geom_bar(data = size_age_R_45_59, aes(x = factor(R), y = prop, fill = sec_cases_grp), stat = "identity") +
  labs(x = expression(R[0]), y = "", title = "Age group index: 45-59",
       fill = "number of cases") +
  scale_fill_manual(values = c("[0]" = "red",
                                 "[1]" = "orange",
                                 "[2, 4]" = "yellow",
                                 "[5, 9]" = "lightgreen",
                                 "[10, 19]" = "green",
                                 "[20, 49]" = "blue",
                                 "[>=50]" = "purple")) +
  scale_y_continuous(breaks = seq(0, 1, 0.25)) +
  theme_minimal()+
  theme(legend.position = "none") + 
  theme(title = element_text(size = 8),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y = element_text(size = 8))  +
  coord_cartesian(ylim = c(0, 1), clip = "off")

plot_45_59 <- plot_45_59_base +
  geom_text(data = label_df,
            aes(x = R, y = prop, label = R),
            size = 2.8,
            inherit.aes = FALSE) + 
  theme(
  axis.title.x = element_text(size = 10, margin = margin(t = 10))
)

###

size_age_R_60_plus <- size_age_R %>%
  filter(age_index_case == "60+")

levels <- c("[0]", "[1]", "[2, 4]", "[5, 9]", "[10, 19]", "[20, 49]", "[>=50]")
  
size_age_R_60_plus$sec_cases_grp <- factor(size_age_R_60_plus$sec_cases_grp, levels = levels)

plot_60_plus_base <- ggplot() +
  geom_bar(data = size_age_R_60_plus, aes(x = factor(R), y = prop, fill = sec_cases_grp), stat = "identity") +
  labs(x = expression(R[0]), y = "", title = "Age group index: 60+",
       fill = "Outbreak size \n(excluding index case)") +
  scale_fill_manual(values = c("[0]" = "red",
                                 "[1]" = "orange",
                                 "[2, 4]" = "yellow",
                                 "[5, 9]" = "lightgreen",
                                 "[10, 19]" = "green",
                                 "[20, 49]" = "blue",
                                 "[>=50]" = "purple")) +
  scale_y_continuous(breaks = seq(0, 1, 0.25)) +
  theme_minimal() + 
  theme(legend.position = "none") +
  theme(title = element_text(size = 8),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y = element_text(size = 8))  +
  coord_cartesian(ylim = c(0, 1), clip = "off")

plot_60_plus <- plot_60_plus_base +
  geom_text(data = label_df,
            aes(x = R, y = prop, label = R),
            size = 2.8,
            inherit.aes = FALSE) + 
  theme(
  axis.title.x = element_text(size = 10, margin = margin(t = 10))
)

```

Figure 7

```{r}
fig_7 <- plot_20_44 + plot_45_59 + plot_60_plus
```

```{r}
sfig_3 <- ggarrange(fig_all_age, fig_7,
                    labels = c("A", "B"),
                    ncol = 1,
                    heights = c(1.5, 1.5))
```

```{r}
ggsave("/Users/df23614/Library/CloudStorage/OneDrive-UniversityofBristol/Shared Documents - grp-HPRU-RT/Avian influenza/Contact patterns/Figure_formatted/SFig_3.tiff", sfig_3, dpi = 300, width = 7.5, height = 4.8, units = "in", device = "tiff", compression = "lzw")
```

# Figure 8 (age distribution of cases)

Data preparation for plotting

Add R column to indicate R_0

```{r}
res_04_0_0_500_0[[2]]$R <- 0
res_04_0_01_500_0[[2]]$R <- 0.1
res_04_0_02_500_0[[2]]$R <- 0.2
res_04_0_03_500_0[[2]]$R <- 0.3
res_04_0_04_500_0[[2]]$R <- 0.4
res_04_0_05_500_0[[2]]$R <- 0.5
res_04_0_06_500_0[[2]]$R <- 0.6
res_04_0_07_500_0[[2]]$R <- 0.7
res_04_0_08_500_0[[2]]$R <- 0.8
res_04_0_09_500_0[[2]]$R <- 0.9
res_04_0_1_500_0[[2]]$R <- 1
res_04_0_11_500_0[[2]]$R <- 1.1
```

Create the data frame for plotting

```{r}
list_no_control <- list(res_04_0_0_500_0, res_04_0_01_500_0, res_04_0_02_500_0, res_04_0_03_500_0,
                        res_04_0_04_500_0, res_04_0_05_500_0, res_04_0_06_500_0, res_04_0_07_500_0,
                        res_04_0_08_500_0, res_04_0_09_500_0, res_04_0_1_500_0, res_04_0_11_500_0)

final_list <- list()

age_grp_index <- c("20-44", "45-59", "60+")

for (i in 1:length(list_no_control)) {
  
  data <- list_no_control[[i]][[2]]
  r_0 <- unique(data$R)
  
  for (age in age_grp_index) {
    
   data_age <- data %>%
     filter(age_index_case == age)
   
   data_clean <- data_age %>%
     filter(secondary_cases != 0)
   
   if (nrow(data_clean) == 0) {
     next
   }
   
   data_clean$prop_0_19 <- data_clean$sec_cases_0_19 / data_clean$secondary_cases
   data_clean$prop_20_44 <- data_clean$sec_cases_20_44 / data_clean$secondary_cases
   data_clean$prop_45_59 <- data_clean$sec_cases_45_59 / data_clean$secondary_cases
   data_clean$prop_60_plus <- data_clean$sec_cases_60_plus / data_clean$secondary_cases
   
   df_age <- data.frame("R_0" = rep(r_0, 4),
                        "age_group_index" = rep(age, 4),
                        "age_group_cases" = c("0-19", "20-44", "45-59", "60+"),
                        "mean_prop" = c(mean(data_clean$prop_0_19), mean(data_clean$prop_20_44),
                                        mean(data_clean$prop_45_59), mean(data_clean$prop_60_plus)))
  
   final_list[[length(final_list) + 1]] <- df_age
   
  }
}

df_temp_0 <- data.frame("R_0" = rep(0, 12),
                      "age_group_index" = c("20-44", "20-44", "20-44", "20-44",
                                            "45-59", "45-59", "45-59", "45-59",
                                            "60+", "60+", "60+", "60+"),
                      "age_group_cases" = c("0-19", "20-44", "45-59", "60+", "0-19", "20-44", "45-59", "60+", "0-19", "20-44", "45-59", "60+"),
                      "mean_prop" = rep(0, 12)
                      )

df_temp_01 <- data.frame("R_0" = rep(0.1, 12),
                      "age_group_index" = c("20-44", "20-44", "20-44", "20-44",
                                            "45-59", "45-59", "45-59", "45-59",
                                            "60+", "60+", "60+", "60+"),
                      "age_group_cases" = c("0-19", "20-44", "45-59", "60+", "0-19", "20-44", "45-59", "60+", "0-19", "20-44", "45-59", "60+"),
                      "mean_prop" = rep(0, 12)
                      )

df_temp <- rbind(df_temp_0, df_temp_01)

final_list[[1]] <- rbind(df_temp, final_list[[1]])


final_data <- do.call(rbind, final_list)

final_data <- final_data %>% mutate(age_group_cases = case_when(
  age_group_cases == "0-19" ~ "0-18",
  age_group_cases == "20-44" ~ "19-44",
  age_group_cases == "45-59" ~ "45-59",
  age_group_cases == "60+" ~ "60+",
))
```

Filter the final dataset by age group of the index case

```{r}
final_data_20_44 <- final_data %>%
  filter(age_group_index == "20-44")

age_dist_20_44_base <- ggplot(final_data_20_44, aes(x = factor(R_0), y = mean_prop, fill = age_group_cases)) +
  geom_bar(stat = "identity") +
  labs(x = expression(R[0]), y = "Proportion of the cases",
       title = "Age index case: 19-44",
       fill = "age group of the cases") +
  theme_minimal()+
  theme(legend.position = "none")  +
  theme(title = element_text(size = 8),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y = element_text(size = 8))  +
  coord_cartesian(ylim = c(0, 1), clip = "off")

age_dist_20_44 <- age_dist_20_44_base +
  geom_text(data = label_df,
            aes(x = R, y = prop, label = R),
            size = 2.5,
            inherit.aes = FALSE) + 
  theme(
  axis.title.x = element_text(size = 10, margin = margin(t = 10))
)

final_data_45_59 <- final_data %>%
  filter(age_group_index == "45-59")

age_dist_45_59_base <- ggplot(final_data_45_59, aes(x = factor(R_0), y = mean_prop, fill = age_group_cases)) +
  geom_bar(stat = "identity") +
  labs(x = expression(R[0]), y = "",
       title = "Age index case: 45-59",
       fill = "age group of the cases") +
  theme_minimal()+
  theme(legend.position = "none")  +
  theme(title = element_text(size = 8),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y = element_text(size = 8))  +
  coord_cartesian(ylim = c(0, 1), clip = "off")

age_dist_45_59 <- age_dist_45_59_base +
  geom_text(data = label_df,
            aes(x = R, y = prop, label = R),
            size = 2.5,
            inherit.aes = FALSE) + 
  theme(
  axis.title.x = element_text(size = 10, margin = margin(t = 10))
)

final_data_60_plus <- final_data %>%
  filter(age_group_index == "60+")

age_dist_60_plus_base <- ggplot(final_data_60_plus, aes(x = factor(R_0), y = mean_prop, fill = age_group_cases)) +
  geom_bar(stat = "identity") +
  labs(x = expression(R[0]), y = "Proportion of the cases",
       title = "Age index case: 60+",
       fill = "Age group of the cases") +
  theme_minimal() +
  theme(title = element_text(size = 8),
        legend.title = element_text(size = 10),
        legend.box.margin = margin(0, 0, 0, -10),
        legend.text = element_text(size = 9),
        legend.key.size = unit(0.8, "lines"),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y = element_text(size = 8))  +
  coord_cartesian(ylim = c(0, 1), clip = "off")

age_dist_60_plus <- age_dist_60_plus_base+
  geom_text(data = label_df,
            aes(x = R, y = prop, label = R),
            size = 2.5,
            inherit.aes = FALSE) + 
  theme(
  axis.title.x = element_text(size = 10, margin = margin(t = 10))
)
```

Figure 8

```{r}
fig_8 <- age_dist_20_44 + age_dist_45_59 + age_dist_60_plus
```

```{r}
ggsave("/Users/df23614/Library/CloudStorage/OneDrive-UniversityofBristol/Shared Documents - grp-HPRU-RT/Avian influenza/Contact patterns/Figure_formatted/SFig_4.tiff", fig_8, dpi = 300, width = 7.5, height = 3.8, units = "in", device = "tiff", compression = "lzw")
```