---
title: "model"
author: "izel"
date: "2025-05-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This code creates the model.

# Load packages

```{r warning=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
library(patchwork) # combine plots with arithmetic operations
library(reshape2)
library(fitdistrplus)
```

# Create the dataframe

```{r}
n_sim <- 3000
```

```{r}
folder_path <- "/Users/df23614/Library/CloudStorage/OneDrive-UniversityofBristol/Shared Documents - grp-HPRU-RT/Avian influenza/Contact patterns/github_codes/AI_spillover_events/outbreak_datasets"
```

## Create infector id, infectee id, age of infector, age of infectee and time of infection

Fix the age group of the index case

``` {r}
outbreak_fn <- function(prob_matrix, size_list, mu_list, contact_list, max_time, R_0, age_groups, index_age_group) {
  
  
  # Initialise the results dataframe
  results <- data.frame(infector = integer(),
                        infectee = integer(),
                        repr_no = integer(),
                        ti = integer(),
                        time = integer(),
                        infector_age = character(),
                        infectee_age = character())
  
  current_infected <- data.frame(infector = 1, infectee = 1, time = 0, infector_age = index_age_group, infectee_age = index_age_group)
  next_infectee <- 2 # Start tracking IDs for new infectees
  
  while (nrow(current_infected) > 0) { # Loop until no new infections occur
    new_infected <- data.frame(infector = integer(),
                               infectee = integer(),
                               repr_no = integer(),
                               ti = integer(),
                               time = integer(),
                               infector_age = character(),
                               infectee_age = character())
    
    for (i in 1:nrow(current_infected)) {
      # Get the infector's age group
      infector_age <- current_infected$infector_age[i]
      
      infector_age_index <- match(infector_age, age_groups)
      
      # Evaluate number of secondary cases for any individual
      infectee_age_group <- age_groups  # Possible infectee age groups
      
      num_sec_cases <- round(rnbinom(1, size = size_list[[infector_age_index]], mu = mu_list[[infector_age_index]]) * R_0 / 6.07)
      
    
        
      if (num_sec_cases > 0) {
        for (l in 1:num_sec_cases) {
          # Draw a time of infection
          time_infection <- round(rgamma(1, shape = 13, rate = 2.0)-3)
          overall_time <- current_infected$time[i] + time_infection
          age_infectee <- sample(age_groups, 1, prob = prob_matrix[infector_age_index, ])
            
          if (overall_time <= max_time) {
            # Add the new infection to new_infected
            new_infected <- rbind(new_infected,
                                    data.frame(infector = current_infected$infectee[i],
                                               infectee = next_infectee,
                                               repr_no = num_sec_cases,
                                               ti = time_infection,
                                               time = overall_time,
                                               infector_age = current_infected$infectee_age[i],
                                               infectee_age = age_infectee))
            next_infectee <- next_infectee + 1 # Increment the infectee ID
          }
        }
      } else {
          # If no infection occurred, keep the infector's time the same
          new_infected <- rbind(new_infected,
                                data.frame(infector = current_infected$infectee[i],
                                           infectee = NA,
                                           repr_no = 0,
                                           ti = NA,
                                           time = current_infected$time[i],  # No change in time
                                           infector_age = current_infected$infectee_age[i],
                                           infectee_age = NA))
          }
      }
    
      # Update results
      results <- rbind(results, new_infected)
    
      # Update current_infected for the next loop
      current_infected <- new_infected[!is.na(new_infected$infectee), ]
  }
  
  return(results)
}
```

## Simulation function for the above model

```{r warning=FALSE}
simulation_fn <- function(prob_matrix, size_list, mu_list, contact_list, max_time , R_0, age_groups, n_sim = n_sim, index_age_group) {
  all_sim_results <- list()
  
  for (sim in 1:n_sim) {
    sim_results <- outbreak_fn(prob_matrix, size_list, mu_list, contact_list, max_time, R_0, age_groups, index_age_group)
    
    # create a new column to track simulations
    sim_results$sim <- sim
  
    all_sim_results[[sim]] <- sim_results
  }
  
  # combine all simulation results rowwise into a single data frame
  outbreak_df <- do.call(rbind, all_sim_results)
  
  return(outbreak_df)
}

```

```{r}
age_groups <- c("0-19", "20-44", "45-59", "60+")

prob_matrix <- matrix(c(0.67, 0.21, 0.075, 0.045,
                        0.133, 0.58, 0.194, 0.093,
                        0.15, 0.40, 0.30, 0.15,
                        0.08, 0.39, 0.31, 0.22), nrow = 4, byrow = TRUE,
                      dimnames = list(age_groups, age_groups))

size_list <- list(1.49, 3.65, 4.46, 4.22)

mu_list <- list(7.85, 4.74, 4.90, 4.17)

contact_list <- list(7.85, 4.74, 4.90, 4.17)

```

## Add symptom status to infector and infectee

```{r warning=FALSE}
symptom_fn <- function(outbreak_df, p_asym) {
  
  outbreak_df$symp_infector <- character(length(outbreak_df$infector))
  outbreak_df$symp_infectee <- character(length(outbreak_df$infector))
  
  for (sim in 1:n_sim) {
    
    # Filter the data for the current simulation
    sim_data <- outbreak_df[outbreak_df$sim == sim, ]
  
    # Loop through each unique infector in the current simulation
    infector_id <- unique(sim_data$infector)
    
    for (id in infector_id) {
      
      # Generate a random value for each row where the infector matches
      symp_status <- runif(1)  
    
      # Assign the corresponding symptom status
      sim_data$symp_infector[sim_data$infector == id] <- ifelse(symp_status > p_asym, "S", "A")
      
    }
    
    infectee_id <- unique(sim_data$infectee)
    
    for (iid in infectee_id) {
      
      if (is.na(iid)) {
      
        sim_data$symp_infectee[which(sim_data$infectee == iid)] <- "NA" 
      } 
      
      if (iid %in% infector_id) {
        
        sim_data$symp_infectee[which(sim_data$infectee == iid)] <- sim_data$symp_infector[which(sim_data$infector == iid)]
        
      } else {
      
        sim_data$symp_infectee[which(sim_data$infectee == iid)] <- ifelse(runif(1) > p_asym, "S", "A")
      }
      
    }
    outbreak_df[outbreak_df$sim == sim, ] <- sim_data
  }
  
  return(outbreak_df)
}
```

## Add tracing status to outbreak_df

```{r warning=FALSE}
tracing_fn <- function(outbreak_df, p_trace) {
  
  outbreak_df$ct_infectee <- character(length(outbreak_df$infector))
  outbreak_df$ct_infector <- character(length(outbreak_df$infector))

  for (sim in 1:n_sim) {
    
    sim_data <- outbreak_df[outbreak_df$sim == sim, ]
  
    # Loop through each unique infector in the current simulation
    infector_id <- unique(sim_data$infector)
  
    for (id in infector_id) {
      
      # Loop through the rows where the infector is equal to the current id
      for (i in which(sim_data$infector == id)) {
        
        # Generate a random value for each occurrence of the infector in the simulation
        contact_tracing <- runif(1)  # Random value for each row
      
        # Assign the corresponding contact tracing status
        if (sim_data$repr_no[i] == 0) {
          sim_data$ct_infectee[i] <- "Not traced"
        } else if (sim_data$symp_infector[i] == "A") {
          sim_data$ct_infectee[i] <- "Not traced"
        } else if (contact_tracing > (1 - p_trace)) {
          sim_data$ct_infectee[i] <- "Traced"
        } else {
          sim_data$ct_infectee[i] <- "Not traced"
        }
        
      } 
      
      if (id == 1) {
        sim_data$ct_infector[which(sim_data$infector == id)] <- "Not traced"
      } else {
      
        sim_data$ct_infector[which(sim_data$infector == id)] <- sim_data$ct_infectee[which(sim_data$infectee == id)]
      
      }
      
    }
    
    outbreak_df[outbreak_df$sim == sim, ] <- sim_data
    
  }
  return(outbreak_df)
}
```

## Add isolation status (whether people would isolate or not)

```{r}
isolation_status_fn <- function(outbreak_df, p_isolate) {
  
  outbreak_df$isolate_infector <- character(length(outbreak_df$infector))
  
  for (sim in 1:n_sim) {
    
    sim_data <- outbreak_df[outbreak_df$sim == sim, ]
    
    infector_id <- unique(sim_data$infector)
    
    for (id in infector_id) {
      
      isolation_status <- runif(1)
      
      symptom_status <- sim_data$symp_infector[which(sim_data$infector == id)]
      
      sim_data$isolate_infector[which(sim_data$infector == id)] <- ifelse(symptom_status == "S" & isolation_status < p_isolate, "isolate", "not_isolate") 
      
    }
    outbreak_df[outbreak_df$sim == sim, ] <- sim_data
  }
  return(outbreak_df)
}

```

## Would infections occur under contact tracing and isolation

```{r warning=FALSE}
effect_tracing_fn <- function(outbreak_df, time_isolation) {
  
  outbreak_df$infected_ct <- "no"

  for (sim in unique(outbreak_df$sim)) {
  
    sim_data <- outbreak_df[outbreak_df$sim == sim, ]
  
    # Loop through all infectees of infector ID = 1
    infected_id_1 <- sim_data$infectee[which(sim_data$infector == 1)]
  
    for (iid in infected_id_1) {
    
      if (is.na(iid)) next  # Skip NA values
    
        # Get all rows where infector is 1
        infector_rows <- which(sim_data$infector == 1)
    
        for (inf_row in infector_rows) {
          ti_value <- sim_data$ti[inf_row]
          isolation_status <- sim_data$isolate_infector[inf_row]
      
          # Apply conditions for infected_ct
          if (isolation_status == "isolate" & ti_value < time_isolation) {
            sim_data$infected_ct[inf_row] <- "yes"
          } else if (isolation_status == "isolate" & ti_value >= time_isolation) {
            sim_data$infected_ct[inf_row] <- "no"
          } else if (isolation_status == "not_isolate") {
            sim_data$infected_ct[inf_row] <- "yes"
          }
        }
    }
  
    # Process infectee IDs to determine their impact as infectors
    infectee_id <- unique(sim_data$infectee)
  
    for (id in infectee_id) {
    
      if (is.na(id)) next  # Skip NA values
    
        # Check if the current infectee has "no" in infected_ct
      if (sim_data$infected_ct[which(sim_data$infectee == id)] == "no") {
        # If yes, ensure no further infections
        sim_data$infected_ct[which(sim_data$infector == id)] <- "no"
      } else {
        # Process conditions for infectees acting as infectors
        infector_rows <- which(sim_data$infector == id)
        
      
        for (inf_row in infector_rows) {
        
          if (is.na(sim_data$infectee[inf_row])) next
        
            ti_value <- sim_data$ti[inf_row]
            symp_status <- sim_data$symp_infector[inf_row]
            ct_status <- sim_data$ct_infector[inf_row]
            isolation_status <- sim_data$isolate_infector[inf_row]
        
          if (symp_status == "S" & isolation_status == "isolate" & ct_status == "Traced" & ti_value > 0) { # traced contacts isolate on the day of onset of symptoms
            sim_data$infected_ct[inf_row] <- "no"
          } else if (symp_status == "S" & isolation_status == "isolate" & ct_status == "Traced" & ti_value <= 0) { # they can infect people before showing symptoms
            sim_data$infected_ct[inf_row] <- "yes"
          } else if (symp_status == "S" & isolation_status == "not_isolate" & ct_status == "Traced" & ti_value > 0) {
            sim_data$infected_ct[inf_row] <- "no"
          } else if (symp_status == "S" & isolation_status == "not_isolate" & ct_status == "Traced" & ti_value <= 0) {
            sim_data$infected_ct[inf_row] <- "yes"
          } else if (symp_status == "S" & isolation_status == "isolate" & ct_status == "Not traced" & ti_value < time_isolation) { # not traced people can voluntarily isolate on day 3 (free parameter)
            sim_data$infected_ct[inf_row] <- "yes"
          } else if (symp_status == "S" & isolation_status == "isolate" & ct_status == "Not traced" & ti_value >= time_isolation) {
            sim_data$infected_ct[inf_row] <- "no"
          } else if (symp_status == "S" & isolation_status == "not_isolate" & ct_status == "Not traced" & ti_value < time_isolation) {
            sim_data$infected_ct[inf_row] <- "yes"
          } else if (symp_status == "S" & isolation_status == "not_isolate" & ct_status == "Not traced" & ti_value >= time_isolation) {
            sim_data$infected_ct[inf_row] <- "yes"
          } else if (symp_status == "A") {
            sim_data$infected_ct[inf_row] <- "yes"
          }
        }
      }
    }
  
    # Update the dataframe
    outbreak_df[outbreak_df$sim == sim, ] <- sim_data
  }
  
  # To reorder the columns so it looks better
  outbreak_df <- outbreak_df[, c(8, 1, 2, 3, 6, 7, 9, 10, 12, 11, 13, 4, 5, 14)]
  
  return(outbreak_df)
}
```

# Analysis of model

## Number of secondary cases and age group of those cases

```{r}
outbreak_size <- function(outbreak_df) {
  
  outbreak_size <- data.frame(sim = 1:n_sim,
                              secondary_cases = rep(NA, n_sim),
                              age_index_case = rep(NA, n_sim),
                              sec_cases_0_19 = rep(NA, n_sim),
                              sec_cases_20_44 = rep(NA, n_sim),
                              sec_cases_45_59 = rep(NA, n_sim),
                              sec_cases_60_plus = rep(NA, n_sim))

 for (sim in unique(outbreak_df$sim)) {
   
   sim_data <- outbreak_df[outbreak_df$sim == sim, ]
   
   # id of the last infectee ignore NA
   last_infectee <- max(sim_data$infectee, na.rm = TRUE)
   
   # if there were no secondary cases, last infectee would be -Inf, for those cases we are asigning the value 0
   outbreak_size$secondary_cases[outbreak_size$sim == sim] <- ifelse(last_infectee == -Inf, 0, last_infectee - 1) # subtract 1 because we are excluding the index case (looking at secondary cases, not everyone)
   
   # for age groups
   table_age_group <- table(sim_data$infectee_age)
   
   outbreak_size$sec_cases_0_19[outbreak_size$sim == sim] <- ifelse(is.na(table_age_group["0-19"]), 0, table_age_group["0-19"])
   
   outbreak_size$sec_cases_20_44[outbreak_size$sim == sim] <- ifelse(is.na(table_age_group["20-44"]), 0, table_age_group["20-44"])
   
   outbreak_size$sec_cases_45_59[outbreak_size$sim == sim] <- ifelse(is.na(table_age_group["45-59"]), 0, table_age_group["45-59"])
   
   outbreak_size$sec_cases_60_plus[outbreak_size$sim == sim] <- ifelse(is.na(table_age_group["60+"]), 0, table_age_group["60+"])
   
   outbreak_size$age_index_case[outbreak_size$sim == sim] <- sim_data$infector_age[which(sim_data$infector == 1)]
 }
 return(outbreak_size)
}
```

## To compare the number of cases before and after contact tracing

```{r warning=FALSE}
compare_fn <- function(outbreak_df) {
  
  compare_ct <- data.frame(sim = 1:n_sim,
                         cases_without_control = rep(NA, n_sim),
                         cases_with_control = rep(NA, n_sim))

  for (sim in unique(outbreak_df$sim)) {
  
    sim_data <- outbreak_df[outbreak_df$sim == sim,]
    
    max_value <- max(sim_data$infectee, na.rm = TRUE)
    
    compare_ct$cases_without_control[compare_ct$sim == sim] <- ifelse(max_value == -Inf, 0, (max_value - 1))
  
    table_ct <- table(sim_data$infected_ct)
  
    if (is.na(table_ct[2])) { 
      table_ct[2] <- 0 # if there are no yes answers, it appears as NA, so assign the value 0
      } 
  
    compare_ct$cases_with_control[compare_ct$sim == sim] <- table_ct[2]  
  
  }

  # Calculate number and the percentage of prevented cases

  for (i in 1:n_sim) {
    
    # percentage of prevented cases
    compare_ct$prevented_cases_perc[i] <- round(100*(compare_ct$cases_without_control[i] - compare_ct$cases_with_control[i]) / (compare_ct$cases_without_control[i]), 1)
    
    # number of prevented cases
    compare_ct$prevented_cases[i] <- compare_ct$cases_without_control[i] - compare_ct$cases_with_control[i]
  
  }
  
  return(compare_ct)
}
```

## Master function for the above model

```{r warning=FALSE}
function_age <- function(p_asym, p_trace, R_0, time_isolation, index_age_group, p_isolate) {
  set.seed(123)
  
  results <- outbreak_fn(prob_matrix = prob_matrix, size_list = size_list, mu_list = mu_list, contact_list = contact_list, max_time = 100, R_0 = R_0, age_groups = age_groups, index_age_group = index_age_group)
  
  
  outbreak_df <- simulation_fn(prob_matrix = prob_matrix, size_list = size_list, mu_list = mu_list, contact_list = contact_list, max_time = 100, R_0 = R_0, age_groups = age_groups, n_sim = n_sim, index_age_group = index_age_group)
  
  outbreak_df <- symptom_fn(outbreak_df, p_asym = p_asym)
  
  outbreak_df <- tracing_fn(outbreak_df, p_trace = p_trace)
  
  outbreak_df <- isolation_status_fn(outbreak_df, p_isolate = p_isolate)
  
  outbreak_df <- effect_tracing_fn(outbreak_df, time_isolation = time_isolation)
  
  outbreak_size <- outbreak_size(outbreak_df)
  
  compare_ct <- compare_fn(outbreak_df)
  
  
  final_results <- list(
    outbreak_df = outbreak_df,
    outbreak_size = outbreak_size,
    compare_ct = compare_ct
  )
  
  return(final_results)
}
```

## Combine data frames across all age groups of index cases

```{r}
combination_function <- function(list_1, list_2, list_3) {
  final_results_2 <- list()
  list_1_new <- list(list_1$outbreak_df, list_2$outbreak_df, list_3$outbreak_df)
  list_2_new <- list(list_1$outbreak_size, list_2$outbreak_size, list_3$outbreak_size)
  list_3_new <- list(list_1$compare_ct, list_2$compare_ct, list_3$compare_ct)
  outbreak_df <- do.call(rbind, list_1_new)
  outbreak_size <- do.call(rbind, list_2_new)
  compare_ct <- do.call(rbind, list_3_new)
  
  final_results_2 <- list(outbreak_df, outbreak_size, compare_ct)
  
  return(final_results_2)
}
```